The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2020-01-26 pkg/partykit/inst/guideliketest/DAGStat2019/slides.Rnw
adapted slides to new structure of simulation results and simulation functions; added slides for continuously changing parameters and full factorial analysis
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2965 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: af9c8b553e4b3bcf72a03ad9d1f029444d4cfbb3

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
set.seed(7)
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/DAGStat2019/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("scales")
library("colorspace")
#source("../guidelike.R")
source("../tests_guideprune.R")
#source("../ccprune.R")

load("../draft/stump.rda")
sim_stump <- simres$res
rm(simres)

load("../draft/tree.rda")
sim_tree <- simres$res
rm(simres)

load("../draft/pruning.rda")
pr <- simres$res
## extract data for xyplot comparing ARI and ARI_p
pr <- pr[,c(1,2,7,8,9,12,15)]
## extended data set with postpruning results being represented as separated tests
prex <- pr[,c(1,2,3,4,6)]
prex <- rbind(prex,prex)

prex$test <- factor(prex$test,
                    levels = c("ctree","mfluc",
                               "guide_sum_12", "guide_sum_1_cor",
                               "ctree_p","mfluc_p",
                               "guide_sum_12_p", "guide_sum_1_cor_p"),
                    labels = c("CTree","MOB",
                               "GUIDE_scores", "GUIDE",
                               "CTree_p","MOB_p",
                               "GUIDE_scores_p", "GUIDE_p"))

for(i in c((length(pr$test)+1) : (2*length(pr$test)))){
  if(prex$test[i] == "CTree") prex$test[i] <- "CTree_p"
  if(prex$test[i] == "MOB") prex$test[i] <- "MOB_p"
  if(prex$test[i] == "GUIDE_scores") prex$test[i] <- "GUIDE_scores_p"
  if(prex$test[i] == "GUIDE") prex$test[i] <- "GUIDE_p"
}

prex$nrsubgr[(length(pr$nrsubgr)+1) : (2*length(pr$nrsubgr))] <- pr$nrsubgr_p
prex$ari[(length(pr$ari)+1) : (2*length(pr$ari))] <- pr$ari_p

load("../draft/3way_xi00_delta03.rda")
load("../draft/3way_xi08_delta1.rda")
d00$xi <- 0
d08$xi <- 0.8
d3way <- rbind(d00, d08)
d3way$xi <- factor(d3way$xi, levels = c(0,0.8), labels = c(0,0.8))

## HCL palette
pal_dgp_light <- qualitative_hcl(3, "Set 2")
pal_dgp <- qualitative_hcl(3, "Dark 3")

## HCL palette (light)
pal_results <- qualitative_hcl(4, "Dark 3")

@@

paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))

@@

class(tp) <- "grapcon_generator"
@

@@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.36\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

\minipage{0.36\textwidth}
<<dgp_stump3, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "beta1", beta0 = beta0, delta = delta, xi = xi, sigma = sigma)

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\end{figure}
\end{frame}

@@

\begin{figure}
\setkeys{Gin}{width=0.8\linewidth}
<<scores_stump, fig=TRUE, echo=FALSE, height=5, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree","mfluc",
                                         "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree","mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

@@

par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

library("latticeExtra")
useOuterStrips(xyplot(prop_Tsplit ~ delta | vary_beta + xi, groups = ~ test, 
                      data = subdataxi, 
                      subset = binary_beta == TRUE & binary_regressor == FALSE & 
                        only_intercept == FALSE,# & xi == 0, 
                      type = "l", lty = c(1,1,1,1),
                      scales=list(x=list(at=seq(1,11,2))),
                      key =  list(text = list(levels(subdataxi$test)),
                                  lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal_results, lwd = 1)),
                      layout= c(3,2),
                      index.cond = list(c(2,3,1),c(2,1)),
                      xlab = expression(delta),
                      ylab = expression("Proportion of selecting"~Z[1]),
                      par.settings = par.settings),

               strip =strip.custom(factor.levels = c(expression("vary"~beta[0]~"and"~beta[1]), 
                                                     expression("vary"~beta[0]), 
                                                     expression("vary"~beta[1])),
                                   bg = "gray90"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0),                                                              expression(xi == 0.8)),
                                       bg = "gray90"))

@
\end{figure}

%\begin{frame}%[fragile]
%\frametitle{Simulation 1: Full factorial analysis (vary both)}

%\begin{figure}%[t]
%\setkeys{Gin}{width=0.8\linewidth}
%<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
%xyplot(pval_z1 ~ cat | xi, 
%       groups = interaction(bin, res_scores), 
%       data = d3way, 
%       ylim = c(0,0.5),
%       type = c("a"), 
%       col = c("red", "blue", "red", "blue"),
%       lty = c(1,1,2,2),
%       xlab = "Categorization",
%       ylab = expression(p~"-value of"~Z[1]),
%       #index.cond = list(c(1,2,3,4)),
%       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
%                                              expression(xi==0.8~"and"~delta==1))),
%       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
%                  lines = list(lty = c(1,1,2,2), 
%                               type = "l",
%                               col = c("red", "blue", "red", "blue"))))
%@
%\end{figure}
%\end{frame}

@@

<<true_tree_dgp, eval = TRUE, echo=FALSE, results=hide>>=

# d <- dgp_tree(nobs = 250)
set.seed(7)
d <- dgp_tree(nobs = 250, xi = 0, binary_regressor = FALSE, vary_beta = "all", 
              binary_beta = TRUE, nrsplitvar = 2, sigma = 0.5)

@@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
#paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.4\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "Y", xlab = "X",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\end{center}
\end{figure}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_tree, test %in% c("ctree","mfluc",
                                        "guide_sum_12", "guide_sum_1_cor"))
#subdata <- subset(subdata, xi %in% c(0,0.5,0.8))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = subdata, 
       type = "l",
       lty = c(1,1,1,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                            bg = "gray90"),
       par.settings = par.settings)

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<prune_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

#prex <- subset(prex, xi %in% c(0,0.5,0.8))

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = prex, 
       type = "l", lty = c(1,1,1,1,2,2,2,2),
       key =  list(text = list(levels(prex$test)[1:4]),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                            bg = "gray90"),
       par.settings = par.settings
)
@
\end{figure}

>>>>

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
set.seed(7)
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/DAGStat2019/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("latticeExtra")
library("scales")
library("colorspace")
#source("../simulation/guidelike.R")
source("../simulation/tests_guideprune.R")
#source("../simulation/ccprune.R")

load("../simulation/results/sim_stump.rda")
load("../simulation/results/sim_tree.rda")
load("../simulation/results/sim_3way.rda")

## HCL palette
pal <- qualitative_hcl(5, "Dark 3")
names(pal) <- c("CTree", "MOB", "GUIDE", "GUIDE+scores", "CTree+max")

## HCL palette for illustrations of data generating process
pal_dgp_light <- qualitative_hcl(3, "Dark 3", c1 = 50, l1 = 75)
pal_dgp <- qualitative_hcl(3, "Dark 3")

## HCL palette for tree panels
paltrees <- rgb(c(0.9, 1), c(0.9, 1), c(0.9, 1))

@@

@@

class(tp) <- "grapcon_generator"

@@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.36\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(1, 2)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

\minipage{0.36\textwidth}
<<dgp_stump3, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, changetype= "abrupt",
               variation = "beta1", beta0 = beta0, delta = delta, xi = xi, sigma = sigma)

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, changetype= "abrupt", 
               variation = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\end{figure}
\end{frame}

@@

\begin{figure}
\setkeys{Gin}{width=0.8\linewidth}
<<scores_stump, fig=TRUE, echo=FALSE, height=5, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree","mfluc",
                                         "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree","mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE+scores", "GUIDE"))

@@

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))

useOuterStrips(xyplot(prop_Tsplit ~ delta | variation + xi, groups = ~ test, 
                      data = subdataxi, 
                      subset = changetype == "abrupt", 
                      type = "l", lty = c(1,1,1,1),
                      lwd = 2,
                      scales=list(x=list(at=seq(1,11,2))),
                      key =  list(text = list(levels(subdataxi$test)),
                                  lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
                      layout= c(3,2),
                      index.cond = list(c(2,3,1),c(2,1)),
                      xlab = expression(delta),
                      ylab = expression("Selection probability of"~Z[1]),
                      par.settings = par.settings),

               strip =strip.custom(factor.levels = c(expression("varying"~beta[0]~"and"~beta[1]), 
                                                     expression("varying"~beta[0]), 
                                                     expression("varying"~beta[1])),
                                   bg = "gray90"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0~" (50%)"),                                                              expression(xi == 0.8~" (90%)" )),
                                       bg = "gray90"))

@
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 1: Continuously changing parameters}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<contbeta, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree", "ctree_max","mfluc","guide_sum_12","guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "ctree_max","mfluc","guide_sum_12","guide_sum_1_cor"),
                       labels = c("CTree","CTree+max","MOB","GUIDE+scores","GUIDE"))
par.settings <- list(superpose.symbol = list(col = pal[c("CTree","CTree+max","MOB","GUIDE+scores","GUIDE")], 
                                             fill = pal[c("CTree","CTree+max","MOB","GUIDE+scores","GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree","CTree+max","MOB","GUIDE+scores","GUIDE")]))

xyplot(prop_Tsplit ~ delta | variation, groups = ~ test, 
       data = subdata, 
       subset = changetype == "continuous", 
       scales=list(x=list(at=seq(1,11,2))),
       type = "l", 
       layout= c(3,1),
       lwd = 2,
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1,1), 
                                type = "l", 
                                col = pal[c("CTree","CTree+max","MOB","GUIDE+scores","GUIDE")], 
                                lwd = 2)), 
       xlab = expression(delta),
       ylab = expression("Selection probability of"~Z[1]),
       index.cond = list(c(2,3,1)),
       strip = strip.custom(factor.levels = c(expression("varying"~beta[0]~" and "~beta[1]),
                                              expression("varying"~beta[0]), 
                                              expression("varying"~beta[1])),
                            bg = "gray90"),
       par.settings = par.settings)
@
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 1: Full factorial analysis of building blocks}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = sim_3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c(pal[1], pal[3], pal[1], pal[3]),
       lty = c(1,1,2,2),
       lwd = 2,
       layout = c(2,1),
       scales = list(alternating=FALSE),
       xlab = "Categorization",
       ylab = expression("p-value of"~Z[1]),
       strip = strip.custom(factor.levels = c(expression(xi==0~"(50%)"~" and"~delta==0.3), 
                                              expression(xi==0.8~"(90%)"~" and"~delta==1)),
                            bg = "gray90"),
       key = list(text = list(c("Goodness of fit:", "  residuals", "  scores", "Dichotomization:", "  yes", "  no")),
                  columns = 2,
                  lines = list(lty = c(0,1,2,0,1,1), 
                               type = "l", lwd = 2,
                               col = c("black","black","black","black", pal[1], pal[3]))))

@
\end{figure}
\end{frame}

@@

<<true_tree_dgp, eval = TRUE, echo=FALSE, results=hide>>=
# d <- dgp_tree(nobs = 250)
set.seed(7)
d <- dgp_tree(nobs = 250, xi = 0, variation = "all", 
              changetype= "abrupt", sigma = 0.5)

@@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
#paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.4\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "Y", xlab = "X",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\end{center}
\end{figure}
\end{frame}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))
xyplot(ari ~ delta | xi, groups = ~ test, 
       data = sim_tree[sim_tree$pruning == "pre",], 
       type = "l", lty = c(1,1,1,1),
       lwd = 2,
       key =  list(text = list(levels(sim_tree$test)[1:4]),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       #index.cond = list(c(2,3,1),c(2,1)),
       par.settings = par.settings,
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                            bg = "gray90"))

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<prune_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))

# add new variable that combines test and pruning strategy
sim_tree$test_pruning <- paste(sim_tree$test, sim_tree$pruning, sep="_")
sim_tree$test_pruning <- factor(sim_tree$test_pruning,
                                levels = c("CTree_pre", "MOB_pre", "GUIDE+scores_pre", "GUIDE_pre",
                                           "CTree_post", "MOB_post", "GUIDE+scores_post", "GUIDE_post"),
                                labels = c("CTree_pre", "MOB_pre", "GUIDE+scores_pre", "GUIDE_pre",
                                           "CTree_post", "MOB_post", "GUIDE+scores_post", "GUIDE_post"))
